Ecology and Conservation Modelling Preregistration Template -- DRAFT

### 2020-05-27 analysis/preregistration_template.Rmd
#10<Hannah>feedback
Rest of Hannah's feedback.
Overall, the prereg template looks fantastic! My main points are 

- remove the intro paragraph to sections that have subheading (e.g. no intro paragraph for section 3.4 if there is a 3.4.1.
- I am not sure about the bit about corroborating on a different dataset in addition to validation described in section 5

### 1.3.2 - Identify outcomes to evaluate decisions under scenarios

*The outcomes should speak directly to the analytical objectives identified in 1.3.1. Outcomes could be qualitative or quantitative, and include descriptive statistic measures, satisficing measures or regret measures (Moallemi et al. 2019). For example, a model outcome in the top 10th percentile of worst conditions, or some minimum performance threshold for meeting a performance objective, or choosing a decision with the minimum regret or least amount of uncertainty.*
To be honest, I don't know how I might apply this to my work or any future works. Maybe I do not have enough knowledge of this type of research to apply it to my basic ecology research. But this isn't to say it's not important, I just wonder if there's a way to make this accessible to non-desicion or modelling peeps. Like, how would I apply it to the woodland condition analysis?

## 1.4 Define candidate actions decisions

*Candidate decisions should be investigated and are specified a priori. Depending on the modelling context, they may be specified by stakeholders, model users or the analyst[@Moallemi2019]. Describe the method used to identify relevant management actions.*

@@

*Describe what processes you will use to elicit and identify relevant scenarios, e.g. literature review, structured workshops with stakeholders or decision-makers. Specify scenarios under which decisions are investigated. Scenarios should be set a priori (i.e. before the model is built, [@Moallemi2019]) and may be stakeholder-defined or expert judgment-driven [@Mahmoud2009].*

@@

*Conceptual models underpin the formal or quantitative model (Cartwright et al. 2016). The conceptual model describes the biological mechanisms relevant to the ecological problem and should capture basic premises about how the target system works, including any prior knowledge and assumptions about system processes. Conceptual models may be representeed in a variety of formats, such as influence diagrams, linguistic model block diagram or bond graphs, and these illustrate how model drivers are linked to both outputs or observed responses, and internal (state) variables (Jakeman, Letcher, and Norton, 2006).*

@@

*Describe what method you will use to elicit or identify the conceptual model. Some common methods include interviews, drawings, and mapping techniques including influence diagrams, cognitive maps and Bayesian belief networks (Moon et al. 2019). (Libby, to provide link to any structured expert elicitation methods?).While it is difficult to decide and justify which method is most appropriate, however Moon et al. (2019) provide guidance addressing this methodological question. Finally, how do you intend on representing the final conceptual model? This will likely depend on the method chosen to elicit the conceptual model.*

## 2.2 Specify key assumptions uncertainties

*This step should list and explain the critical conceptual design decisions, including: spatial and temporal scales, selection of entities and processes, representation of stochasticity and heterogeneity, consideration of local versus global interactions, environmental drivers, etc. [@Grimm:2014es]. The influence of particular theories, concepts, or importantly, earlier models, should be explained and justified against alternative conceptual design decisions that might lead to alternative model structures [@Grimm:2014es]. Specify key assumptions and uncertainties underlying the model's design. Describe how uncertainty and variation will be represented in this model. This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].*

## 2.3 Define conceptual model: identify predictor response variables

@@

@@

*Collect, process and prepare data available for parameterisation, determining model structure, and for scenario analysis.*

@@

*Document the identity, quantity and provenance of any data used to develop, identify and test the model. Describe how the data is arranged, in terms of replication and covariates.Explain how you will summarise what data and knowledge will be used to design and parameterise the model, including references, data sources, and information about whom, when and where those data were collected [@Grimm:2014es].*

**For pre-existing data: -- as appropriate.**

- *For each dataset, is the data open or publically available? (Y/N)*
- *How can the data be accessed? Provide a link or contact as appropriate, indicating any restrctions on the use of data.*
- *Date of download, access, or future access:*
- *Describe the source of the data - what entity originally collected this data? (National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
- *Codebook. If a Codebook is available, please link to it here and / or upload the document.*
- *Prior work based on this dataset - Have you published / presented any previous work based on this dataset? Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont his dataset you have worked on.*
- *Prior Research Activity - Have you worked with these data before? Describe any prior but unpublished research activity using these data. Be specific and transparent.*
- *Prior knowledge of the current dataset - Describe any prior knowledge of the dataset. Be specific.*

@@

- Data collection procedures - Please describe your data collection process, including how sites and transects or any other physical unit were selected. Describe any inclusion or exclusion rules, and the study timeline. 
- Sample Size - Describe the sample size of your study.
- Sample Size Rationale - Describe how you determined the appropriate sample size for your study. It could include feasibility constraints, such as time, money or personnel. 
- If sample size cannot be specified, specify a stopping rule - i.e. how will you decide when to terminate your data collection?

@@

*Describe how you will separate and distinguish between raw data, manipulated data, and outputs from modelling or any analyses of the model.*

@@

*In most modelling cases, it is necessary to perform preliminary analyses to understand the data and check that assumptions and requirements of the chosen modelling procedures are met. Data exploration prior to model fitting or development may include exploratory analyses to check for collinearity, spatial and temporal coverage, quality and resolution, outliers, or the need for transformations [@Yates2018]. Describe how you will summarise and explore your data, and explain what method you will use (graphical, tabular or otherwise) to represent your data and any analyses.*

@@

*Describe how you will evaluate the quality and sources of both numerical and qualitative data that available for model parameterisation, and for determining the overall model structure [@Grimm:2014es]. Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009]. Document any issues with data reliability. This is important because data quality and ecological relevance might be constrained by measurement error, inappropriate experimental design, and heterogeneity and variability inherent in ecological systems [@Grimm:2014es].*

@@

*Describe how the model will be critically evaluated. Evaluation includes both the overall model structure, and any submodels if relevant. How will any simplifying assumptiosn be critically assessed[@Augusiak:2014gz]? Explain whether this process will include consultation or feedback from a client, manager, or model user.*

@@

*Critical decisions are made here about the type of model, modelling framework and approach to be used. The model is formalised into a mathematical / statistical description of the system and respective changes it can experience  [@Mahmoud2009]. In this section describe what quantitative methods you will use to build the model/s, explain how they are relevant to the client/manager/user's purpose.*

@@

*Describe what class or approach of model you will use and explain how the choice of model class was informed by the analytical objectives of the model. Modelling approaches lie on a spectrum from correlative or phenomenological to mechanistic or process-based  [@Yates2018]; where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data [@White2019a].*

@@

*All modelling approaches require the selection of model features, which conform with the conceptual model and data specified in previous steps (Jakeman 2006). The choice of model are determined in conjunction with features are selected. Usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller (Jakeman, 2006). However, if changes to these fundamental aspects of the model do need to change, please document how and why these choices were made.*

@@

*Specify how you will operationalise response variables in the model. This should relate directly to the analytical and or management objectives specified during the problem formulation phase. Specifications could include:- qualitative change, such as a direction of change- the extent of a response- an extreme value- a trend- a long-term mean- a probability distribution- a spatial pattern- a time-series, or- the frequency, location, or probability of some event occuring.Provide a rationale for your choices, including why plausible alternatives were not chosen (Jakeman, 2006).*

@@

*Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions. Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase. Include in your rationalse for selection detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.*

@@

*Specify which types of variables are covered in the model, and the nature of their treatment (e.g. lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006). Specify model structural features, including:- the functional form of interactions,- data structures- measures used to specify links, spatio temporal scales and processes as well as their interactions- any bins or discretisation of continuous variables.*

@@

*This refers to the iterative process of determining the most efficient or parsimonious representation of the system at the appropriate scale of concern (Jakeman, 2006) that best meets the analytical objectives specified in the problem formulation phase. Approaches to finding model structure and parameters may be knowledge-supported, or data-driven [@Boets:2015gl]. Model selection methods can include traditional inferential approaches such as unconstrained searches of a dataset for patterns that explain variations in the response variable, or use of ensemble-modelling methods (Barnard et al. 2019). Ensemble modelling procedures might aim to derive a single model, or a multi-model average [@Yates2018]. Refining actions to develop a model could include iteratively dropping parameters or adding them, or aggregating / disaggregating system descriptors, such as dimensionality and processes (Jakeman, 2006). Specify what approach and methods you will use to identify model structure and parameters.*

@@

*Before calibrating the model to the data, the performance criteria on which the calibration is judged are chosen. These criteria and their underlying assumptions should reflect the desired properties of the parameter estimates / structure [@Jakeman:2006ii]. For example, modellers might seek that parameter estimates are robust to outliers, unbiased, and yield appropriate predictive performance. Modellers will need to consider whether the assumptions of the estimation technique yielding those desired criteria are appropriate to the problem at hand. For integrated or sub-divided models, other considerations might include choices about where to disaggregate the model for parameter estimation; e.g. spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii]. Specifying performance criteria a priori is important because it involves pre-specifying how the fitted or quantitative model will be interpreted in advance, avoiding biases like confirmation bias, HARKing, and cherry-picking some performance tests from the full suite of tests undertaken.This is important - because it involves pre-specifying how we will interpret the fitted or quantitative model to avoid cognitive biases like confirmation bias, HARKing or other questionable research practices.*

### 3.4.1 Choice of estimation performance criteria

@@

### 3.4.2 Choice of structure and parameter estimation technique

*Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g. distributed boundary conditions). For example, will you calibrate all variables simultaneously by optimising fi of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006). Identify which variables were parameterised directly, such as by expert elicitation or prior knowledge.Specify which algorithm(s) you will use for any data-driven structure / parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].*

@@

*Once critical decisions have been made about the approach and method of model specification, translate the conceptual model into the quantitative model. For data-driven and model-selection approaches that determine model structure and parameters, describe any initial model specifications and parameterisations, including for any tune-in parameters. (Should this go here or in model calibration?)*

@@

*This phase involves fitting the formally defined model to data and checking and testing the fitted model (what should go in thisphase if using a numerical or mathematical, but not correlative model or how could we change the wording to make this optional if not relevant?)*

@@

*Describe the validation scheme you will use for model testing and evaluation. Please explain your reasoning for your choice of model calibration and validation scheme. The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation)  [@Yates2018;[@Barnard2019]. Although using independent data is the gold-standard or idealised practice for model evaluation, Typically testing on independent datasets is not possible for large and/or integrated models, especially when they are being used to generate anticipatory predictions beyond the conditions on which they were calibrated [@Jakeman:2006ii]. Describe and justify the method for your calibration and validation schema.*

## 4.1.2 Data partitioning

*Describe the approach for partitioning the data in support of the calibration and validation schema described above. For example, specify the number of folds that will be created and the relative size of each.*

*How will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced?*

@@

*What Quality Assurance measures will you take to verify the modle has been correctly implemented as per the formal model specification ? This could be as simple as an informal ‘model walkthrough' to examine the model's behaviour qualitatively, or it could include more sophisticated approaches such as formal unit testing approaches using continuous integration platforms (see [@Yenni2019]). Qualitative tests could include syntax checking of code, and code reviews by peers. Checks for verification implementation should include i) thoroughy checking for bugs or programming errors,  and ii) whether the implemented model performs as dictated by the model description  [@Grimm:2014es]. Specifying up front QA tests for implementation verification may help to avoid selective debugging.*

@@

*"Model Checking" goes by many names ("conditional verification", "quantitative verification", "model output verification" <cite>), and refers to a series of analyses that assess a model's performance in representing the system of interest [@Conn:2018hd]. During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences. Model checking therefore heps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system [@Conn:2018hd]. Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].*

### 4.3.1 Quantitative tests

*Specify any diagnostics or tests you will use during model checking. For each test, specify the numerical threshold or criteria that will you use to interpret the outcome of the test in assessing the model's ability to sufficiently represent the gathered data used to develop and parameterise the model.*

@@

*This step is largely informal and case-specific, but requires ‘face validation' with model users / clients / managers who aren't involved in the development of the model to assess whether the interactions and outcomes of the model are feasible an ddefensible  [@Grimm:2014es]. This process could be called a “laugh test" or a “pub test, and in addition to checking the model's believability, it builds the client's confidence in the model [@Jakeman:2006ii]. Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment. Briefly explain how you will qualitatively check the model, and whether + how you will include users and clients in the process.*

# 5.0 Model Validation and Evaluation

*Model Validation & Evaluation - This phase consists of a suite of analyses that collectively inform the decision and whether and when a model is suitable to meet its intended purpose (Augusiak, Van den Brink and Grimm 2014). Errors in design and implementation of the model and their implication on the model output are assessed. Ideally independent data is used against the model outputs to assess whether the model output behaviour exhibits the required accuracy for the model's intended purpose. The outcomes of these analyses build confidence in the model applications and increase understandingof model strengths and limitations. Model evaluation including model analysis and other testing should complement model checking. It should evaluate model checking, and consider over-fitting and extrapolation. The higher the proportion of calibrated, or uncertain parameters, the greater the risk that the model seems to work correctly […] but for the wrong reasons" Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mecahnisms of the system (Grimm). Evaluation of model outputs against external data in conjunction with the results from model checking provide information about the structural realism and therefore credibility of the model (Grimm et al. 2014).*

## 5.1 Model output corroboration

*Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model. Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step <2.4>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames — hence, why ecologists often model in the first place (Grimm, et al. 2014. Independent predictions can instead be tested on submodels. Alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model's output is (Grimm et al.2014). If unable to evaluate the model outputs against independent data, explain why and explain what steps you will take to interrogate the model.*

@@

## 5.3 Evaluate model assumptions

*Explain how you will demonstrate a lack of violation of or robustness to model assumptions. If you cannot perform quantitative tests, specify what theoretical justifications would justify a lack of violation of or robustness to model assumptions. If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.(Complete in follow-up iterations of the pre-registration, only if there are departures from assumptions as demonstrated in the planned tests above). The consequences of assumption violations on the interpretation of results should be assessed [@Araujo2019]. Specify how you will assess departures from assumptions. If assumption violations cannot be avoided, explain describe how you will explore their consequences.*

## 5.4 Model analysis

>>>>

### 1.3.2 - Identify how the model will tell you about your analytic objectives

*The outcomes should speak directly to the analytical objectives identified in 1.3.1. Outcomes could be qualitative or quantitative. For example, you might be using your model in a scenario analysis to determine which management decision is associated with minimum regret or the highest likelihood of improvement.*
To be honest, I don't know how I might apply this to my work or any future works. Maybe I do not have enough knowledge of this type of research to apply it to my basic ecology research. But this isn't to say it's not important, I just wonder if there's a way to make this accessible to non-desicion or modelling peeps. Like, how would I apply it to the woodland condition analysis? 
HF - Made the changes above in response to the comment.... do they help? If not, I'm struggling to understand exactly what this step is achieving.

## 1.4 Define candidate decisions and method of identifying relevant management actions

*Candidate decisions should be investigated and are specified a priori. Depending on the modelling context, they may be specified by stakeholders, model users or the analyst[@Moallemi2019].¬†Describe the method used to identify relevant management actions.*

@@

*Describe what processes you will use to elicit and identify relevant scenarios, e.g. literature review, structured workshops with stakeholders or decision-makers. Specify scenarios under which decisions are investigated. Scenarios should be set a priori (i.e. before the model is built, [@Moallemi2019]) and may be stakeholder-defined or driven by the judgement of the modeller or other experts [@Mahmoud2009].*

@@

*Conceptual models underpin the formal or quantitative model (Cartwright et al. 2016). The conceptual model describes the biological mechanisms relevant to the ecological problem and should capture basic premises about how the target system works, including any prior knowledge and assumptions about system processes. Conceptual models may be representeed in a variety of formats, such as influence diagrams, linguistic model block diagram or bond graphs, and these illustrate how model drivers are linked to both outputs or observed responses, and internal (state) variables (Jakeman, Letcher, and Norton, 2006).*
HF- Love this step

@@

*Describe what method you will use to elicit or identify the conceptual model. Some common methods include interviews, drawings, and mapping techniques including influence diagrams, cognitive maps and Bayesian belief networks (Moon et al. 2019). (Libby, to provide link to any structured expert elicitation methods?).While it is difficult to decide and justify which method is most appropriate, however Moon et al. (2019) provide guidance addressing this methodological question. Finally, how do you intend on representing the final conceptual model? This will likely depend on the method chosen to elicit the conceptual model.*
HF- not sure if this step is required - I think we should expect people to have already done this bit before they start their preregistration... Or we could put it into 2.0 as an addendum - "if you haven't defined your conceptual model yet describe how you plan to do so."

## 2.2 Specify key assumptions and uncertainties

*This step should list and explain the critical conceptual design decisions, including: spatial and temporal scales, selection of entities and processes, representation of stochasticity and heterogeneity, consideration of local versus global interactions, environmental drivers, etc. [@Grimm:2014es]. The influence of particular theories, concepts, or importantly, earlier models, should be explained and justified against alternative conceptual design decisions that might lead to alternative model structures [@Grimm:2014es].*

HF comment: the above part of 2.2 seems like a big ask. I think we want to achieve that we will need to break it down further. Would it be sufficient to just have the bit below?

*Specify key assumptions and uncertainties underlying the model's design. Describe how uncertainty and variation will be represented in this model. This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].*

## 2.3 Identify predictor and response variables

@@

HF- love this step

@@

*Collect, process and prepare data available for parameterisation, determining model structure, and for scenario analysis.*
HF - this isn't super clear to me. Maybe we need more clarification here or an example

@@

*Document the identity, quantity and provenance of any data that will be used to develop, identify and test the model. Describe how the data is arranged, in terms of replicates and covariates.*

**For pre-existing data: -- delete as appropriate.**

- *For each dataset, is the data open or publically available? (Y/N)*
- *How can the data be accessed? Provide a link or contact as appropriate, indicating any restrctions on the use of data.*
- *Date of download, access, or future access:*
- *Describe the source of the data - what entity originally collected this data? (National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
- *Codebook and meta-data. If a Codebook or other meta-data is available, please link to it here and / or upload the document(s).*
- *Prior work based on this dataset - Have you published / presented any previous work based on this dataset? Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont this dataset you have worked on.* HF- Love this question
- *Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.*
- *Prior knowledge of the current dataset - Describe any prior knowledge of the dataset. Be specific.* HF- not sure how someone would answer this question. Maybe we need a more specific Q or an example

@@

- Data collection procedures - Please describe your data collection process, including how sites and transects or any other physical unit were selected and arranged. Describe any inclusion or exclusion rules, and the study timeline. 
- Sample Size - Describe the sample size of your study.
- Sample Size Rationale - Describe how you determined the appropriate sample size for your study. It could include feasibility constraints, such as time, money or personnel. 
- If sample size cannot be specified, specify a stopping rule - i.e. how will you decide when to terminate your data collection?

@@

*Describe how you will separate and distinguish between raw data, manipulated data, and outputs from modelling or any analyses of the model.* HF - Not sure exactly what this means

@@

*In most modelling cases, it is necessary to perform preliminary analyses to understand the data and check that assumptions and requirements of the chosen modelling procedures are met. Data exploration prior to model fitting or development may include exploratory analyses to check for collinearity, spatial and temporal coverage, quality and resolution, outliers, or the need for transformations [@Yates2018]. Describe how you will summarise and explore your data, and explain what method you will use (graphical, tabular or otherwise) to represent your data and any analyses.*
HF- should we include something here asking people to specify what they intend to find out from the data exploration?

@@

*Describe how you will determine how reliable the data is for the given purpose. Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].*

*Document any issues with data reliability. This is important because data quality and ecological relevance might be constrained by measurement error, inappropriate experimental design, and heterogeneity and variability inherent in ecological systems [@Grimm:2014es].*

@@

*Describe how your conceptual model will be critically evaluated. Evaluation includes both the completeness and suitability of the overall model structure.*

*How will any simplifying assumptions be critically assessed [@Augusiak:2014gz]?* HF- I'm not sure how to answer this one. Maybe an example would help?

*†Explain whether this process will include consultation or feedback from a client, manager, or model user.*

@@

*In this section describe what quantitative methods you will use to build the model/s, explain how they are relevant to the client/manager/user's purpose.*

@@

*Describe what class or approach of model you will use and¬†explain how the choice of model class was informed by the analytical objectives of the model. Modelling approaches lie on a spectrum from correlative or phenomenological to mechanistic or process-based¬† [@Yates2018]; where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data [@White2019a].*

@@

*All modelling approaches require the selection of model features, which conform with the conceptual model and data specified in previous steps (Jakeman 2006). The choice of model are determined in conjunction with features are selected. Usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller (Jakeman, 2006). However, if changes to these fundamental aspects of the model do need to change, please document how and why these choices were made.*
HF- it's not clear to me what a model feature is and I find this a little confusing. I wonder if we need the text describing the 3.2 heading when it's stepped out in the subsections.

@@

*Specify how you will operationalise the response variables in the model. This should relate directly to the analytical and or management objectives specified during the problem formulation phase. Operationalisations could include:- the extent of a response- an extreme value- a trend- a long-term mean- a probability distribution- a spatial pattern- a time-series, - qualitative change, such as a direction of change or- the frequency, location, or probability of some event occuring. Provide a rationale for your choices, including why plausible alternatives were not chosen (Jakeman, 2006).*
HF- this is great. Only slight concern is the deceptively difficult task that last sentence represents

@@

*Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions. Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase.* HF- great

*Include in your rational for selection, detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.* HF - I feel like this belongs somewhere else. I'm not sure where thinking about which variables the model outputs are sensitive to would relate to me choosing between logistic and cloglog models etc

@@

*Specify which variables are included in the model, and the nature of their treatment (e.g. lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006).*

*Specify model structural features, including:- the functional form of interactions,- data structures- measures used to specify links,- any bins or discretisation of continuous variables.*

@@

*This refers to the process of determining the best/most efficient/parsimonious representation of the system at the appropriate scale of concern (Jakeman, 2006) that best meets the analytical objectives specified in the problem formulation phase. Approaches to finding model structure and parameters may be knowledge-supported, or data-driven [@Boets:2015gl]. Model selection methods can include traditional inferential approaches such as unconstrained searches of a dataset for patterns that explain variations in the response variable, or use of ensemble-modelling methods (Barnard et al. 2019). Ensemble modelling procedures might aim to derive a single model, or a multi-model average [@Yates2018]. Refining actions to develop a model could include iteratively dropping parameters or adding them, or aggregating / disaggregating system descriptors, such as dimensionality and processes (Jakeman, 2006). Specify what approach and methods you will use to identify model structure and parameters.*

@@

*Before calibrating the model to the data, the performance criteria on which the calibration is judged are chosen. These criteria and their underlying assumptions should reflect the desired properties of the parameter estimates / structure [@Jakeman:2006ii]. For example, modellers might seek parameter estimates that are robust to outliers, unbiased, and yield appropriate predictive performance. Modellers will need to consider whether the assumptions of the estimation technique yielding those desired criteria are appropriate to the problem at hand. For integrated or sub-divided models, other considerations might include choices about where to disaggregate the model for parameter estimation; e.g. spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii]. Specifying performance criteria a priori is important because it involves pre-specifying how the fitted or quantitative model will be interpreted in advance, avoiding biases like confirmation bias, HARKing, and cherry-picking some performance tests from the full suite of tests undertaken.*
HF- I think this paragraph is only relevant to performance criteria- doesn't really explore estimation technique. As above though, I'm not sure that this paragraph is neessary given that it's stepped out in the subheadings below

### 3.4.1 Choice of performance criteria

@@

### 3.4.2 Choice of parameter estimation technique	
HF- I think that it might be double dipping to include structure here as well as in 3.3
*Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g. distributed boundary conditions). For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006). Identify which variables were parameterised directly, such as by expert elicitation or prior knowledge. Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].*

@@

*Once critical decisions have been made about the approach and method of model specification,¬†translate the conceptual model into the quantitative model. For data-driven and model-selection approaches that determine model structure and parameters, describe any initial model specifications and parameterisations, including for any tune-in parameters. (Should this go here or in model calibration?)*
HF- I thought that this had already been covered in 3.2 and 3.3. Help me understand how this is distinct?

@@

*This phase involves fitting the formally defined model to data and checking and testing the fitted model (what should go in thisphase if using a numerical or mathematical, but not correlative model or how could we change the wording to make this optional if not relevant?)*
HF- I dont think this paragraph is necessary

@@

*Describe how you will validate and check the calibration of the model. The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation)  [@Yates2018;[@Barnard2019]. Justify your choice*

## 4.1.2 Describe validation data

*If partitioning data for cross-validation: describe the approach specifying the number of folds that will be created and the relative size of each. Describe how will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced.*

*If using external data: describe any known differences between the training and validation datasets and the size of the validation dataset. Describe how will you document and/or share the external data such that any subsequent modelling can be computationally reproduced.*

@@

*What Quality Assurance measures will you take to verify the model has been correctly implemented? Tests could include syntax checking of code,¬†and code reviews by peers. Checks for verification implementation should include i) thoroughy checking for bugs or programming errors, ¬†and ii) whether the implemented model performs as dictated by the model description  [@Grimm:2014es]. Specifying up front quality assurance tests for implementation verification may help to avoid selective debugging.*
HF- I've tweaked this section to make it more distinct from 4.3

@@

*"Model Checking" goes by many names ("conditional verification", "quantitative verification", "model output verification" <cite>), and refers to a series of analyses that assess a model's performance in representing the system of interest [@Conn:2018hd].  Model checking therefore heps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system [@Conn:2018hd]. Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].*
HF- I've oved some of the information otu of 4.3 and put it in the sub headings so that this paragraph isn't necessary

### 4.3.1 Quantitative Model Checking

*Specify any diagnostics or tests you will use during model checking to assess a model's performance in representing the system of interest. During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences. For each test, specify the criteria that will you use to interpret the outcome of the test in assessing the model's ability to sufficiently represent the gathered data used to develop and parameterise the model.*

@@

*This step is largely informal and case-specific, but requires ‚Äòface validation' with model users / clients / managers who aren't involved in the development of the model to assess whether the interactions and outcomes of the model are feasible an ddefensible  [@Grimm:2014es]. This process could be called a ‚Äúlaugh test" or a ‚Äúpub test, and in addition to checking the model's believability, it builds the client's confidence in the model [@Jakeman:2006ii].¬†Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment. Briefly explain how you will qualitatively check the model, and whether + how you will include users and clients in the process.*
HF- great section

# 5.0 Model Validation and Evaluation
HF- I think that this is covered in section 4 with descriptions of validation and checking. I'm not sure how this section is different.

*Model Validation & Evaluation - This phase consists of a suite of analyses that collectively inform the decision and whether and when a model is suitable to meet its intended purpose (Augusiak, Van den Brink and Grimm 2014). Errors in design and implementation of the model and their implication on the model output are assessed. Ideally independent data is used against the model outputs to assess whether the model output behaviour exhibits the required accuracy for the model's intended purpose. The outcomes of these analyses build confidence in the model applications and increase understandingof model strengths and limitations. Model evaluation including model analysis and other testing should complement model checking. It should evaluate model checking, and consider over-fitting and extrapolation. The higher the proportion of calibrated, or uncertain parameters, the greater the risk that the model seems to work correctly [‚Ä¶] but for the wrong reasons" Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mecahnisms of the system (Grimm). Evaluation of model outputs against external data in conjunction with the results from model checking provide information about the structural realism and therefore credibility of the model (Grimm et al. 2014).*

## 5.1 Model output corroboration	
HF- This sounds like a second round of validation. I dont think I've ever seen this happen.
*Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model. Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset¬†minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step <2.4>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames ‚Äî hence, why ecologists often model in the first place (Grimm, et al. 2014. Independent predictions can instead be tested on submodels. Alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model's output is (Grimm et al.2014). If unable to evaluate the model outputs against independent data, explain why and explain what steps you will take to interrogate the model.*

@@

## 5.3 Evaluate model assumptions	
HF- I would put this up in section 4 and get rid of section 5
*Explain how you will demonstrate a lack of violation of or robustness to model assumptions. If you cannot perform quantitative tests, specify what theoretical justifications would justify a lack of violation of or robustness to model assumptions. If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.(Complete in follow-up iterations of the pre-registration, only if there are departures from assumptions as demonstrated in the planned tests above). The consequences of assumption violations on the interpretation of results should be assessed [@Araujo2019]. Specify how you will assess departures from assumptions. If assumption violations cannot be avoided, explain describe how you will explore their consequences.*
